CODECHECK: An open-science initiative for the independent execution of computations underlying research articles.

### 2021-01-11 codecheck-paper.tex
more acks confirmed

If adopted today, a CODECHECK system, even if temporal in a sustainable 
transition towards more open publication and review practices, can contribute
to increased trust in the results of research. Introducing CODECHECK
should be informed by lessons learned from (introducing) open peer review
(OPR) \cite{ross-hellauer_guidelines_2019}.
All our conversations with publishers and editors at some point
touched on the procedural and technical implications of openness, be it 
that OPR is neither applied nor well-known by stakeholders, or that process
management software would not support such open communication.
We would be keen to use the flexibility of the principles and cooperate
on non-public CODECHECKs to learn more about the advantages and yet unclear
specific challenges. Does CODECHECK work better with OPR?
Each rollout of CODECHECK should be accompanied by studies to ensure 
codechecking does not fall into the same gaps as peer review did \cite{tennant_limitations_2020}
and to ensure positive change within the respective peer review system.
More reproducible practices initiated by CODECHECKs could lead
communities/disciplines to reach a state where authors do provide good
enough material and reviewers have acquired broady enough skills that
the latter will generally conduct a CODECHECK-level of checking, and
only in especially sophisticated cases will a specialised codechecker
be included.
The main challenge for us remains to get journals to embrace the
idea behind CODECHECK and realise processes conform to the principles,
whether they use the actual name for it or not.
The CODECHECK badge is, intentionally, not branded and simply states
``code works''.
This \emph{cultural change}, however, is needed for the valuation of the
efforts that go into proper evaluation of papers.
Multiple adopting journals will also help to answer relevant questions on
the establishing process (What are crucial decisions or pain points? Can
authors retract code/data once a CODECHECK has started?),
what variants of CODECHECKs will be most common, and how open CODECHECKs
may influence the scope and anonymity of conventional review over time.

@@

% question mark indicates we are waiting for person to respond.
We are grateful to the following individuals for discussions regarding
the work presented here: Andy Collings, Melissa Harrison?, Giuliano
Maciocci, Naomi Penfold? (eLife), Rebecca Kirk (PLOS Computational Biology), Scott
Edmunds (GigaScience), and Andrew Hufton? (Scientific Data). Iain Davies and
Yuhao (Sebastian) Wang developed code and example certificates.
We thank Antonio P\'{a}ez? (Journal of Geographical Systems) for enabling
CODECHECKs, Carlos Granell and Frank Ostermann? for contributing
certificates as reproducibility reviewers at the AGILE conference, and 
all authors of auditable workflows for their participation.

>>>>

If adopted today, a CODECHECK system, even if temporal in a sustainable 
transition towards more open publication and review practices, can contribute
to increased trust in the results of research. Introducing CODECHECK
should be informed by lessons learned from (introducing) open peer review
(OPR) \cite{ross-hellauer_guidelines_2019}.
All our conversations with publishers and editors at some point
touched on the procedural and technical implications of openness, be it 
that OPR is neither applied nor well-known by stakeholders, or that process
management software would not support such open communication.
We would be keen to use the flexibility of the principles and cooperate
on non-public CODECHECKs to learn more about the advantages and yet unclear
specific challenges. Does CODECHECK work better with OPR?
Each rollout of CODECHECK should be accompanied by studies to ensure 
codechecking does not fall into the same gaps as peer review did \cite{tennant_limitations_2020}
and to ensure positive change within the respective peer review system.
More reproducible practices initiated by CODECHECKs could lead
communities/disciplines to reach a state where authors do provide good
enough material and reviewers have acquired broady enough skills that
the latter will generally conduct a CODECHECK-level of checking, and
only in especially sophisticated cases will a specialised codechecker
be included.
The main challenge for us remains to get journals to embrace the
idea behind CODECHECK and realise processes conform to the principles,
whether they use the actual name for it or not.
The CODECHECK badge is, intentionally, not branded beyond the checkmark
and green colour and simply states ``code works''.
This \emph{cultural change}, however, is needed for the valuation of the
efforts that go into proper evaluation of papers.
Multiple adopting journals will also help to answer relevant questions on
the establishing process (What are crucial decisions or pain points? Can
authors retract code/data once a CODECHECK has started?),
what variants of CODECHECKs will be most common, and how open CODECHECKs
may influence the scope and anonymity of conventional review over time.

@@

% question mark indicates we are waiting for person to respond.
We are grateful to the following individuals for discussions regarding
the work presented here: Andy Collings, Melissa Harrison?, Giuliano
Maciocci, Naomi Penfold (eLife), Rebecca Kirk (PLOS Computational Biology), Scott
Edmunds (GigaScience), and Andrew Hufton? (Scientific Data). Iain Davies and
Yuhao (Sebastian) Wang developed code and example certificates.
We thank Antonio P\'{a}ez (Journal of Geographical Systems) for enabling
CODECHECKs, Carlos Granell and Frank Ostermann for contributing
certificates as reproducibility reviewers at the AGILE conference, and 
all authors of auditable workflows for their participation.
